Improving inference of evolutionary history

### 2014-11-03 slides/osu/slides.tex
Updating osu slides.

\begin{document}
\maketitle

% I. What I do
% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
%   a. Dynamic planet---need to account for large-scale processes
%       i. E.g., SE Asia---dynamic landscape
\begin{frame}
    \frametitle{Processes of diversification}
        \begin{itemize}
            \item<1-> Large-scale geological and climatic processes are
                important in biodiversification and community assembly
            \item<2-> Accounting for such processes will better our
                understanding of biodiversity
            \item<3-> We need methods for inferring evolutionary patterns
                predicted by historical events from contemporary populations
        \end{itemize}
\end{frame}

@@

        \begin{uncoverenv}<2->
        \colorbox{white}{
            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
                \begin{onlyenv}<2-3>
                \begin{itemize}
                    \item<2-> Repeated coalescence and fragmentation of island complexes
                    \item<3-> Prominent paradigm for explaining Philippine biodiversity
                \end{itemize}
                \end{onlyenv}

@@

\ignore{
\begin{frame}
\begin{columns}[c]
    \column{.5\textwidth}
        \begin{table}%[htbp]
            \scriptsize
            \addtolength{\tabcolsep}{-0.09cm}
            \centering
            \begin{tabular}{ l c c }
                % \toprule
                \textbf{Species} & {\boldmath $n_1$} & {\boldmath $n_2$} \\
                \hline
                \textbf{\textsc{Mammals}} & & \\
                \emph{Crocidura beatus}             & 12 & 11 \\
                \emph{Crocidura negrina-panayensis} & 12 & 6  \\
                \emph{Hipposideros obscurus}        & 19 & 9  \\
                \emph{Hipposideros pygmaeus}        & 3  & 12 \\
                \emph{Cynopterus brachyotis}        & 20 & 8  \\
                \emph{Cynopterus brachyotis}        & 8  & 14 \\
                \emph{Haplonycteris fischeri}       & 29 & 8  \\
                \emph{Haplonycteris fischeri}       & 9  & 21 \\
                \emph{Macroglossus minimus}         & 19 & 4  \\
                \emph{Macroglossus minimus}         & 8  & 10 \\
                \emph{Ptenochirus jagori}           & 4  & 7  \\
                \emph{Ptenochirus jagori}           & 8  & 8  \\
                \emph{Ptenochirus minor}            & 30 & 9  \\
                \textbf{\textsc{Squamates}} & & \\
                \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
                \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
                \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
                \emph{Gekko mindorensis}            & 8  & 11 \\
                \emph{Insulasaurus arborens}        & 22 & 10 \\
                \emph{Pinoyscincus jagori}          & 8  & 8  \\
                \emph{Dendrelaphis marenae}         & 6  & 6  \\
                \textbf{\textsc{Anurans}}  & & \\
                \emph{Limnonectes leytensis}        & 4  & 2  \\
                \emph{Limnonectes magnus}           & 2  & 3  \\
                \hline
            \end{tabular}
        \end{table}
    \column{.5\textwidth}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=1.5cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/gekko-mindorensis.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=1.5cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1>[height=1.5cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
\end{columns}
\end{frame}

\begin{frame}
\begin{columns}[c]
    \column{.5\textwidth}
        \begin{table}%[htbp]
            \scriptsize
            \addtolength{\tabcolsep}{-0.09cm}
            \def\firstrowcolor{}
            % \def\secondrowcolor{}
            % \def\thirdrowcolor{}
            % \def\fourthrowcolor{}
            % \def\fifthrowcolor{}
            \only<2>{\def\firstrowcolor{\rowcolor{rowred}}}
            % \only<3>{\def\secondrowcolor{\rowcolor{rowred}}}
            % \only<4>{\def\thirdrowcolor{\rowcolor{rowred}}}
            % \only<5>{\def\fourthrowcolor{\rowcolor{rowred}}}
            % \only<6>{\def\fifthrowcolor{\rowcolor{rowred}}}
            \centering
            \begin{tabular}{ l c c }
                % \toprule
                \textbf{Species} & {\boldmath $n_1$} & {\boldmath $n_2$} \\
                \hline
                \textbf{\textsc{Mammals}} & & \\
                  \emph{Crocidura beatus}             & 12 & 11 \\
                \firstrowcolor  \emph{Crocidura negrina-panayensis} & 12 & 6  \\
                  \emph{Hipposideros obscurus}        & 19 & 9  \\
                  \emph{Hipposideros pygmaeus}        & 3  & 12 \\
                  \emph{Cynopterus brachyotis}        & 20 & 8  \\
                \firstrowcolor  \emph{Cynopterus brachyotis}        & 8  & 14 \\
                  \emph{Haplonycteris fischeri}       & 29 & 8  \\
                \firstrowcolor  \emph{Haplonycteris fischeri}       & 9  & 21 \\
                  \emph{Macroglossus minimus}         & 19 & 4  \\
                \firstrowcolor  \emph{Macroglossus minimus}         & 8  & 10 \\
                  \emph{Ptenochirus jagori}           & 4  & 7  \\
                \firstrowcolor  \emph{Ptenochirus jagori}           & 8  & 8  \\
                  \emph{Ptenochirus minor}            & 30 & 9  \\
                \textbf{\textsc{Squamates}} & & \\
                  \emph{Cyrtodactylus gubaot-sumuroi} & 29 & 6  \\
                  \emph{Cyrtodactylus annulatus}      & 14 & 3  \\
                \firstrowcolor  \emph{Cyrtodactylus philippinicus}  & 6  & 14 \\
                \firstrowcolor  \emph{Gekko mindorensis}            & 8  & 11 \\
                \firstrowcolor  \emph{Insulasaurus arborens}        & 22 & 10 \\
                  \emph{Pinoyscincus jagori}          & 8  & 8  \\
                \firstrowcolor  \emph{Dendrelaphis marenae}         & 6  & 6  \\
                \textbf{\textsc{Anurans}}  & & \\
                  \emph{Limnonectes leytensis}        & 4  & 2  \\
                  \emph{Limnonectes magnus}           & 2  & 3  \\
                \hline
            \end{tabular}
        \end{table}
    \column{.5\textwidth}
        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
        \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
        % \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-bohol_mindanao.png}
        % \includegraphics<4>[width=\textwidth]{../images/maps/Philippines-leyte_mindanao.png}
        % \includegraphics<5>[width=\textwidth]{../images/maps/Philippines-mindanao_samar.png}
        % \includegraphics<6>[width=\textwidth]{../images/maps/Philippines-leyte_samar.png}
\end{columns}
\end{frame}
}

    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \includegraphics<1>[width=\textwidth]{../images/maps/Philippines.png}
        \includegraphics<2>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
        \begin{onlyenv}<3->
            \begin{itemize}
                \item Strong support for simultaneous divergence of all 22 taxon pairs

@@

\begin{frame}
    \frametitle{New method: \dppmsbayes}
    \begin{itemize}[<+->]
        % \item<1-> Reparameterized the model implemented in \msb
        \item<1-> Replaced uniform priors on continuous parameters with gamma and
            beta distributions
        \item<1-> Dirichlet process prior (DPP) over all possible divergence
            models
        \item<1-> Develop multi-processing interface to improve performance for
            accommodating genomic datasets
        % \item<4-> Uniform prior over divergence models
    \end{itemize}
\end{frame}

@@

\begin{frame}
    \frametitle{More data}
    \begin{columns}[c]
        \column{.5\textwidth}
        \begin{myitemize}
            \item Collecting genomic data from two gekkonid radiations
                co-distributed Southeast Asia
            \begin{myitemize}
                \item \emph{Cyrtodactylus}---12+ species
                \item \emph{Gekko}---13+ species
            \end{myitemize}
            \item Whole genome assemblies for each genus
            \item RAD-seq libraries constructed for $\approx150$ individuals
                from each genus
        \end{myitemize}
        \column{.5\textwidth}
        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
    \end{columns}
\end{frame}

@@

\begin{frame}
    \frametitle{More power}
    \begin{itemize}[<+->]
        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
            \begin{itemize}
                \item Uses all the information in the sequence data
                \item Applicable to much deeper timescales
                \item \ldots Currently inefficient
            \end{itemize}
        \item Working to incorporate new developments on analytical integration
            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
            \begin{itemize}
                \item Very efficient numerical approximation of posterior 
                \item Applicable to NGS datasets
            \end{itemize}
    \end{itemize}
\end{frame}

@@

%   b. Empirical
%       i. Gobi (NGS --> NSF)
%       ii. West-Central Africa
%           1. unfunded collaborator
%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
%           1. Planning pre-proposal
%       iv. Long-term study of small island of sunda shelf
%           1. Tiny islands with pops of Eutropis multifasciata and
%           Lepidodactylus lugubris
%           2. Inundated during last interglacial ~120,000 years ago, spent the
%           better part of next 100k years as part of exposed sunda shelf, then
%           fragmented during current interglacial
%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
%           parthenogenesis), skinks in color patterns
%           4. Affects of recent, severe fragmentation on diversification and
%           demography
%           5. Comparative genomics to understand evolution of color and
%           reproductive mode
%           6. Evolutionary ecology---what ecological context influence
%           evolution asexuality/live-bearing/color.
\begin{frame}
    \frametitle{Future directions---Empirical}
    \begin{itemize}
        \item<1-> Many exciting empirical systems for applying new methods!
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Gobi Desert}
    \vspace{-2mm}
    \begin{columns}
        % \column{\dimexpr\paperwidth-10pt}
        \column{\dimexpr\paperwidth}
        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Gobi Desert}
    \vspace{-2mm}
    \begin{columns}
        % \column{\dimexpr\paperwidth-10pt}
        \column{\dimexpr\paperwidth}
        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
    \end{columns}
\end{frame}

\begin{frame}[t]
    \frametitle{West-Central Africa}
    \begin{center}
        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
        \hspace{0.3mm}
        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure4_herps.jpeg}
    \end{center}

@@

\begin{frame}
    \frametitle{Questions?}    
    \begin{center}
        {
        \Large
        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
        }
    \end{center}
\end{frame}

>>>>

\begin{document}
% \maketitle
\begin{frame}
    \begin{columns}[c]
        \column{.5\textwidth}
            \maketitle
        \column{.5\textwidth}
            \begin{figure}
                \begin{center}
                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
                \end{center}
            \end{figure}
    \end{columns}
\end{frame}

% I. What I do
% II. What I've been focusing on recently (started in PhD and continuing in postdoc)
%   a. Dynamic planet---need to account for large-scale processes
%       i. E.g., SE Asia---dynamic landscape
{
\setbeamercolor{background canvas}{bg=black}
\setbeamercolor{whitetext}{fg=white}
\begin{frame}
    \frametitle{Processes of diversification}
    \begin{columns}[c]
        \column{.5\textwidth}
        {\usebeamercolor[fg]{whitetext}
        \begin{itemize}
            \item<2-> Large-scale geological and climatic processes are
                important in biodiversification and community assembly
            \item<3-> Accounting for such processes will better our
                understanding of biodiversity
            \item<4-> We need methods for inferring evolutionary patterns
                predicted by historical events from contemporary populations
        \end{itemize}
        }
        \column{.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
            \end{center}
        \end{figure}
    \end{columns}
\end{frame}
}

@@

        \begin{uncoverenv}<2->
        \colorbox{white}{
            \begin{minipage}[t][3.6cm][t]{1.1\textwidth}
                \begin{onlyenv}<2-3>
                \begin{itemize}
                    \item<2-> Repeated coalescence and fragmentation of island complexes
                    \item<3-> Prominent paradigm for explaining biodiversity
                \end{itemize}
                \end{onlyenv}

@@

\begin{frame}
    \frametitle{Climate-driven diversification}
\begin{columns}[c]
    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{onlyenv}<1>
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/rafe.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/cam.jpg}}
        \centerline{
        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/charles.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=3cm]{../../../../Dropbox/field-photos/people/jake.jpg}}
        \end{minipage}
        \end{onlyenv}

        \begin{onlyenv}<2->
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
        \end{minipage}
        \end{onlyenv}

    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
        \begin{onlyenv}<4->
            \begin{itemize}
                \item Strong support for simultaneous divergence of all 22 taxon pairs

@@

\begin{frame}
    \frametitle{New method: \dppmsbayes}
    \begin{itemize}[<+->]
        % \item<1-> Reparameterized the model implemented in \msb
        \item<1-> Replaced uniform priors on continuous parameters with gamma and
            beta distributions
        \item<2-> Dirichlet process prior (DPP) over all possible divergence
            models
        \item<3-> Develop multi-processing interface to improve performance for
            accommodating genomic datasets
        % \item<4-> Uniform prior over divergence models
    \end{itemize}
\end{frame}

@@

\begin{frame}
    \frametitle{More data}
    \begin{columns}[c]
        \column{.5\textwidth}
        \begin{myitemize}
            \item Collecting genomic data from two gekkonid radiations
                co-distributed Southeast Asia
            \begin{myitemize}
                \item \emph{Cyrtodactylus}---12+ species
                \item \emph{Gekko}---13+ species
            \end{myitemize}
            \item Whole genome assemblies for each genus
            \item RAD-seq libraries constructed for $\approx150$ individuals
                from each genus
        \end{myitemize}
        \centerline{
        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/misc/lee-and-me.jpg}
        \hspace{2mm}
        \includegraphics<1->[height=2.5cm]{../../../../Dropbox/field-photos/people/rafe.jpg}}
        \column{.5\textwidth}
        \includegraphics[width=\textwidth]{../images/photos/cyrt-agusanensis.jpg} \quad
        \includegraphics[width=\textwidth]{../images/photos/gekko-mindorensis.jpg}
    \end{columns}
\end{frame}

@@

\begin{frame}
    \frametitle{More power}
    \centerline{
    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/mtholder.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/jeet2.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{../../../../Dropbox/field-photos/people/vladimir.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{../images/revbayes.png}}

    \begin{itemize}[<+->]
        \item Full-likelihood Bayesian implementation \footnote{\tiny\shortfullcite{JeetDiss}}
            \begin{itemize}
                \item Uses all the information in the sequence data
                \item Applicable to much deeper timescales
                \item \ldots Currently inefficient
            \end{itemize}
        \item Working to incorporate new developments on analytical integration
            of gene trees \footnote{\tiny\shortfullcite{Bryant2012}}
            \begin{itemize}
                \item Very efficient numerical approximation of posterior 
                \item Applicable to NGS datasets
            \end{itemize}
    \end{itemize}
\end{frame}

@@

%   b. Empirical
%       i. Gobi (NGS --> NSF)
%       ii. West-Central Africa
%           1. unfunded collaborator
%       iii. Compare diversification of mainland SE Asia to Oceanic Islands
%           1. Planning pre-proposal
%       iv. Long-term study of small island of sunda shelf
%           1. Tiny islands with pops of Eutropis multifasciata and
%           Lepidodactylus lugubris
%           2. Inundated during last interglacial ~120,000 years ago, spent the
%           better part of next 100k years as part of exposed sunda shelf, then
%           fragmented during current interglacial
%           3. Both species vary in reproductive mode (oviparity/ovoviviparity,
%           parthenogenesis), skinks in color patterns
%           4. Affects of recent, severe fragmentation on diversification and
%           demography
%           5. Comparative genomics to understand evolution of color and
%           reproductive mode
%           6. Evolutionary ecology---what ecological context influence
%           evolution asexuality/live-bearing/color.
{
\setbeamercolor{background canvas}{bg=black}
\setbeamercolor{whitetext}{fg=white}
\begin{frame}
    \frametitle{Processes of diversification}
    \begin{columns}[c]
        \column{.5\textwidth}
        {\usebeamercolor[fg]{whitetext}
        \begin{itemize}
            \item<1-> These comparative tools will allow evolutionary
                biologists to leverage genomic data to infer shared
                evolutionary history
            \item<2-> This will better our understanding of how community scale
                processes affect the generation and assembly of biodiversity
            \item<3-> Many exciting empirical systems for applying new methods!
        \end{itemize}
        }
        \column{.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
            \end{center}
        \end{figure}
    \end{columns}
\end{frame}
}

\begin{frame}
    \frametitle{Future directions: Gobi Desert}
    \vspace{-2mm}
    \begin{columns}
        % \column{\dimexpr\paperwidth-10pt}
        \column{\dimexpr\paperwidth}
        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-auto-shop.jpg}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Future directions: Gobi Desert}
    \vspace{-2mm}
    \begin{columns}
        % \column{\dimexpr\paperwidth-10pt}
        \column{\dimexpr\paperwidth}
        \includegraphics[width=\textwidth]{../../../../Dropbox/field-photos/gobi/mongolia-teratoscincus-przewalskii-landscape.jpg}
    \end{columns}
\end{frame}

\begin{frame}[t]
    \frametitle{Future directions: West-Central Africa}
    \begin{center}
        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure2_WAmapForestDist.png}
        \hspace{0.3mm}
        \includegraphics<1->[height={3cm}]{../../../../Dropbox/field-photos/africa/Figure4_herps.jpeg}
    \end{center}

    \only<1>{
    \begin{center}
        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/adam.png}
        \hspace{0.3mm}
        \includegraphics[height={3.5cm}]{../../../../Dropbox/field-photos/people/fujita.jpg}
    \end{center}
    }

@@

\begin{frame}
    \frametitle{Questions?}    
    \begin{columns}[c]
        \column{.5\textwidth}
        \begin{center}
            {
            \Large
            \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
            }
        \end{center}
        \column{.5\textwidth}
            \begin{figure}
                \begin{center}
                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
                \end{center}
            \end{figure}
    \end{columns}
\end{frame}
